NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all

### 2023-04-14 main.tex

\subsection{Applications of the NMRlipids Databank}
\subsubsection{Selecting simulation parameters using NMRlipids Databank: Best models for most abundant neutral membrane lipids}
%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations. 
%
%For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}. On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}. 
%
To minimize the detrimental consequences of artificial MD simulation results for their applications, the quality of lipid bilayer MD simulations has to be carefully assessed~\cite{antila22b}. This can be done, for example, against the C--H bond order parameters from NMR spectroscopy and the form factors from X-ray scattering~\cite{ollila16}, although it requires comparisons between large number of simulations, which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}. To streamline this process, we have defined quantitative quality measures that enable automatic ranking of lipid bilayer simulations based on their quality against experiments. Conformational ensembles of individual lipid molecules are evaluated in the NMRlipids Databank by calculating the probabilities for each C--H bond order parameter to locate within experimental error ($P^{\mathrm{hg}}$, $P^{\mathrm{sn1}}$,$P^{\mathrm{sn2}}$, and $P^{\mathrm{total}}$). Furthermore, the quality against X-ray scattering experiments ($FF_q$) is estimated as the difference in the experimental and simulated locations of the first form factor minimum. These measures can be used to evaluate membrane dimensions, as the form factor minima locations and acyl chain C--H bond order parameters correlate with the membrane lateral packing and thickness (Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}). In addition, the ergodicity of conformational sampling of lipids is estimated by calculating $\tau_\mathrm{rel}$, the convergence time of the slowest principal component divided by the simulation length. Let us demonstrate how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for simulations of membranes with the two most biologically-abundant neutral membrane lipids~\cite{vanmeer08}: POPC and POPE.

@@

H. S. A. contributed to the quality evaluation metrics based on NMR and SAXS, and was involved in testing the GUI and polishing the manuscript.
Lara S. Bort implemented about 630 pre-existing lipid double layer simulations into the databank layer repository and created corresponding mapping files.
P.B. implemented the PCA-based equilibration metric, supervised the work of A.K., and participated in code refactoring.
B.K. contributed to the initial design and implementation of the databank's Python back-end and was involved with the writing the manuscript.
R.G.-F was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
A.P. was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
A.K. implemented the PCA-based equilibration metric. 
I.G. advised on implementation of the PCA-based equilibration metric and supervised the work of A.K.
N.K. contributed experimental scattering data and SIMtoEXP algorithm.
M.K. provided simulations with OPLS3e force field
J.J.M. provided simulations with Amber force field, general discussion, and manuscript editing

>>>>

\subsection{Applications of the NMRlipids Databank}
\subsubsection{Selecting simulation parameters using NMRlipids Databank: Best models for most abundant neutral membrane lipids}
%Quality of membrane simulations with different force fields have been evaluated against experimental data during parameterization and in separate comparison studies~\cite{botan15,ollila16,catte16,pluhackova16,perez17,leonard19,??}, but the lack of universal quality measure for membrane simulations complicates the estimations of reliability of simulations. 
%
%For example, OPLS3e parameters overcome CHARMM36 in structural quality for POPC but predicts overestimated ion binding~\cite{kurki22}. On the other hand, CHARMM36 gives the best description for lipid headgroup conformational ensembles~\cite{bacle21}, while GROMOS-CKP parameters best capture the membrane packing in POPS bilayers~\cite{antila22b}. 
%
To minimize the detrimental consequences of artificial MD simulation results for their applications, the quality of lipid bilayer MD simulations has to be carefully assessed~\cite{antila22b}. This can be done, for example, against the C--H bond order parameters from NMR spectroscopy~\cite{bacle21,Wur23} and the form factors from X-ray scattering~\cite{ollila16}, although it requires comparisons between large number of simulations, which is laborious even with collaborative approaches~\cite{botan15,catte16,antila19,bacle21}. To streamline this process, we have defined quantitative quality measures that enable automatic ranking of lipid bilayer simulations based on their quality against experiments. Conformational ensembles of individual lipid molecules are evaluated in the NMRlipids Databank by calculating the probabilities for each C--H bond order parameter to locate within experimental error ($P^{\mathrm{hg}}$, $P^{\mathrm{sn1}}$,$P^{\mathrm{sn2}}$, and $P^{\mathrm{total}}$). Furthermore, the quality against X-ray scattering experiments ($FF_q$) is estimated as the difference in the experimental and simulated locations of the first form factor minimum. These measures can be used to evaluate membrane dimensions, as the form factor minima locations and acyl chain C--H bond order parameters correlate with the membrane lateral packing and thickness (Figs.~\ref{fig:quality}G,~\ref{fig:QualityCorrelationsSI}, and~\ref{fig:sizedependence}). In addition, the ergodicity of conformational sampling of lipids is estimated by calculating $\tau_\mathrm{rel}$, the convergence time of the slowest principal component divided by the simulation length. Let us demonstrate how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for simulations of membranes with the two most biologically-abundant neutral membrane lipids~\cite{vanmeer08}: POPC and POPE.

@@

H. S. A. contributed to the quality evaluation metrics based on NMR and SAXS, and was involved in testing the GUI and polishing the manuscript.
Lara S. Bort implemented about 630 pre-existing lipid double layer simulations into the databank layer repository and created corresponding mapping files.
P.B. implemented the PCA-based equilibration metric, supervised the work of A.K., and participated in code refactoring.
B.K. contributed to the initial design and implementation of the databank's Python back-end and was involved with the writing the manuscript.
R.G.-F was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
A.P. was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
A.K. implemented the PCA-based equilibration metric. 
I.G. advised on implementation of the PCA-based equilibration metric and supervised the work of A.K.
N.K. contributed experimental scattering data and SIMtoEXP algorithm.
M.K. provided simulations with OPLS3e force field. 
T.M.F. performed and analysed the solid-state NMR experiments.
